---
title: "Governors: Temperance"
format:
  html: default
execute:
  echo: false
---

```{r}
#| label: setup
#| message: false

library(tidyverse)
library(tidymodels)
library(broom)
library(marginaleffects)

governors <- read_csv("data/governors.csv", show_col_types = FALSE)

# Carried forward from Wisdom: the close-election analysis tibble.
x <- governors |>
  filter(abs(win_margin) <= 5, year > 1945) |>
  mutate(election_result = if_else(win_margin > 0, "Win", "Lose"))

# Carried forward from Courage: the data generating mechanism.
fit_governors <- linear_reg() |>
  fit(lived_after ~ election_result + win_margin + election_age + party, data = x)
```

## The Question

Barfort, Klemmensen, and Larsen (2021) collected the lifespans of 3,587 candidates for US gubernatorial office going back to the 1800s, in order to estimate the causal effect of winning office on longevity. By comparing winners and losers of *close* elections — near coin flips — they argue the treatment is as good as randomly assigned. (See Gelman's commentary on the paper, and the authors' response, for a running argument about whether the headline estimate is believable.)

There are two scenarios. Focus on just one of them:

- Imagine you work for a life insurer, and want to forecast how long candidates for the US Senate might live based on their age, party, election result, and other variables.

- Imagine you are a researcher. You want to know if winning candidates live longer. You are interested in candidates for any elected office --- senate, governor, mayor, et cetera.

The life insurer asks: How many years, on average, will this year's US Senate candidates live after their election, given their age, party, and election result?

The researcher asks: What is the average causal effect of winning an election --- for any elected office, from senator to governor to mayor --- rather than losing it, on the number of years a candidate lives afterward?

Note that we use the same data, and the same fitted model, to help both people.

The setup chunk of this document carries forward the work from Wisdom and Courage: the close-election tibble `x` and the fitted model `fit_governors` are already built, so today starts at the answering.

## Temperance

### Predictions

Use `predictions(fit_governors)` to estimate the years lived after the election for individual candidates in the sample.

```{r}
predictions(fit_governors) |> 
  as_tibble() |>
    select(election_result, election_age, party, conf.low, conf.high) |> 
      head(6)
```


Use `avg_predictions(fit_governors, by = "election_result")` to estimate the average years lived after the election for winners and losers.

```{r}
avg_predictions(fit_governors, by = "election_result")
```

What is the expected number of years lived after the election for a losing candidate (with 95% CI)?

The expected number of years is an average of 27.2 years with a confidence interval between 25.2-29.2.


What is the expected number of years lived after the election for a winning candidate (with 95% CI)?

The expected number of years is an average of 29.6 years with a confidence interval between 27.7-31.5.


### Comparisons

Use `avg_comparisons(fit_governors, variables = "election_result")` to estimate the effect of winning on years lived after the election.

```{r}
avg_comparisons(fit_governors, variables = "election_result")
```

Winning results in an average of living 8.63 years longer (with a confidence interval of 3.26 to 14).


Why is the estimate from `avg_comparisons()` so much larger than the difference between the two group averages from `avg_predictions()`?

avg_predictions averages each group under their own win margins. Meanwhile, avg_comparisons averages within each row while having win_margins fixed.

### Visualizations

Create a visualization of the average predictions by election result using `plot_predictions()`.

```{r}
plot_predictions(fit_governors, by = "election_result", draw = FALSE) |>
  ggplot(aes(x = election_result, y = estimate, ymin = conf.low, ymax = conf.high, color = election_result)) +
  geom_pointrange(size = 1) +
  scale_color_manual(values = c("Lose" = "#d95f02", "Win" = "#1b9e77")) +
  labs(
    title = "Expected Post-Election Lifespan of Gubernatorial Candidates",
    subtitle = "Average predicted years lived after election by outcome",
    caption = "Source: Barfort, Klemmensen, and Larsen (2021)",
    x = "Election Outcome",
    y = "Predicted Years Lived After Election (Years)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "none",
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 6)),
    plot.subtitle = element_text(color = "grey30", size = 11, margin = margin(b = 12)),
    plot.caption = element_text(color = "grey50", size = 8, hjust = 0, margin = margin(t = 12)),
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )
```

What measure of uncertainty is shown in this plot?

The plot shows the 95% confidence intervals.

Create a plot showing predicted years lived after the election as a function of `election_age`, separately for winners and losers.

```{r}
plot_predictions(fit_governors, condition = c("election_age", "election_result"), draw = FALSE) |>
  ggplot(aes(x = election_age, y = estimate, ymin = conf.low, ymax = conf.high, color = election_result, fill = election_result)) +
  geom_ribbon(alpha = 0.2, color = NA) +
  geom_line(linewidth = 1) +
  scale_color_manual(values = c("Lose" = "#d95f02", "Win" = "#1b9e77")) +
  scale_fill_manual(values = c("Lose" = "#d95f02", "Win" = "#1b9e77")) +
  labs(
    title = "Predicted Candidate Longevity by Age and Election Outcome",
    subtitle = "Predicted years lived after election as a function of age",
    caption = "Source: Barfort, Klemmensen, and Larsen (2021)",
    x = "Age at Election (Years)",
    y = "Predicted Years Lived After Election",
    color = "Election Outcome",
    fill = "Election Outcome"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14, margin = margin(b = 6)),
    plot.subtitle = element_text(color = "grey30", size = 11, margin = margin(b = 12)),
    plot.caption = element_text(color = "grey50", size = 8, hjust = 0, margin = margin(t = 12)),
    legend.position = "bottom",
    axis.title.x = element_text(margin = margin(t = 10)),
    axis.title.y = element_text(margin = margin(r = 10))
  )
```

### Interpretation

How does the life insurer interpret the estimated effect of winning? How does the researcher interpret the same number differently?

The life insurer does not really interpret this. They use the DGM to predict per unit life expectancy. Meanwhile, the researcher interprets this as the causal effect.

### Humility

Besides the average years lived, identify two additional quantities of interest the person in your scenario might care about.

Party and wealth.


Why might these estimates be a poor answer to your scenario's question? Identify at least two possible sources of bias.

1) The sample is small, and the confidence intervals are large.
2) The treatment is winning a gubernatorial election, while the life insurer needs to focus on winning Senate elections.
